---
title: "penguins"
author: "Rafael Tobias Miras"
date: "2024-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instalando paquetes y cargando librerias

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```


# Cargamos el df y visualizamos el contenido
```{r}
penguins <- palmerpenguins::penguins
```

```{r}
View(penguins)
glimpse(penguins)
str(penguins)
```


## Gráficos

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g, )
) + geom_point()
```

## Añadimos colores por especies

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) + geom_point()
```

## Añadimos curva que muestre la relación entre body_mass y flipper_length para cada especie

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) + 
        geom_point() +
        geom_smooth(method = "lm")
```

Unimos la lineas en una sola aunque no acaba de estar perfectamente definida.

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
        geom_point(mapping = aes(color = species)) +
        geom_smooth(method = "lm")
```

A parte de indicar un color por especie es conveniente incluir una forma diferente por especie para facilitar la visualización a
personas con dificultades visuales como pueden ser el daltonismo, añadimos "shape = species" en geom_point

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
        geom_point(mapping = aes(color = species, shape = species)) +
        geom_smooth(method = "lm")
```

Le pones título, subtítulo y nombre a los ejes.

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
) + 
        geom_point(mapping = aes(color = species, shape = species)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Body mass and flipper lenght",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                x = "Flipper lenght (mm)" , y = "Body mass (g)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()
```

## EXERCICIES

1. How many rows are in penguins? How many columns?

```{r}
str(penguins)

```
334 rows and 8 columns


2. What does the bill_depth_mm variable in the penguins data frame descrive? Read the help for ?penguins to dins out.

The bill_depth_mm variable descrives bill depth in  millimeters (profundidad del pico in spanish)


3. Make a scatterplot of bill_depth_mm vs bill_lenght_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_lenght_mm on the x-axis. Describe the relationship between the two variables

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = bill_length_mm, y = bill_depth_mm)
) + 
        geom_point(mapping = aes(color = species, shape = species)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Bill lenght and bill depth",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                x = "Bill lenght (mm)" , y = "Bill depth (mm)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()
```
We can see a negative relationship

4. What happens if you make a scatterplot of species versus bill_depth_mm?

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = species, y = bill_depth_mm)
) + 
        geom_point(mapping = aes(color = species, shape = species)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Bill lenght and bill depth",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                x = "species" , y = "Bill depth (mm)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()
```

What might be a better choice of genom?

Boxplot is better to check the result. we can check mean, quartils, outliers, etc. More info with this kind of plot

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = species, y = bill_depth_mm)
) + 
        geom_boxplot(mapping = aes(color = species, shape = species)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Bill lenght and bill depth",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                x = "species" , y = "Bill depth (mm)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()
```


5. Why does the following give an error, and how would you fix it?

```{r}
ggplot(data = penguins) +
        geom_point()
```

The plot requires values for x and y

6. Why does the na.rm argument do in geom_point()? What is the default value of the argument? Create a scatterplot where you succesfully use this arguments ser to TRUE

The default value for the argument na.rm is FALSE. If it's TRUE it'll remove the NA values and with a warning. If it's TRUE it'll remove the NA values silently


```{r}
ggplot(
        data = penguins,
        mapping = aes(x = bill_length_mm, y = bill_depth_mm)
) + 
        geom_point(mapping = aes(color = species, shape = species, na.rm = TRUE)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Bill lenght and bill depth",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                x = "Bill lenght (mm)" , y = "Bill depth (mm)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()
```

7. Add the following caption to the plot you made in the previous excercise: "Data come from the palmerpenguins package". Hint: Take a look at hte documentation for labs()

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = bill_length_mm, y = bill_depth_mm)
) + 
        geom_point(mapping = aes(color = species, shape = species, na.rm = TRUE)) +
        geom_smooth(method = "lm") +
        labs(
                title = "Bill lenght and bill depth",
                subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
                caption = "Data come from the palmerpenguins package",
                x = "Bill lenght (mm)" , y = "Bill depth (mm)" ,
                color = "Species" , shape = "Species"
        ) +
        scale_color_colorblind()

```

8. Re-create the following visualization. What aesthetic should bill_depth_mm be mapped to? And shoud be mapped at global level or at geom level?

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g, fill = bill_depth_mm)
) + 
        geom_point() +
        geom_smooth(method = "lm") +
        labs(
                color = "Bill depth" , shape = "Bill depth"
        ) +
        scale_color_colorblind()
```

9. run this code in your head and predict what the output will look like. then, run the code in R and check your predictions

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
        geom_point() +
        geom_smooth(se = FALSE)
```

10. Will these two graphs look different? Why/why not?

```{r}
ggplot(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
        geom_point() +
        geom_smooth()

```

```{r}
ggplot() +
        geom_point(data = penguins,
                   mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
        geom_smooth(
                data = penguins,
                mapping = aes(x = flipper_length_mm, y = body_mass_g)
        )
```

## VISUALISING DISTRIBUTIONS

## Categorical variable

```{r}
ggplot(penguins, aes(x = species)) + 
geom_bar()
```
We are reordering levels with fct_infreq() function, trnasforming them to factor

```{r}
ggplot(penguins, aes(x = fct_infreq(species))) + 
geom_bar()
```


## Numerical variable

```{r}
ggplot(penguins, aes(x = body_mass_g)) + 
        geom_histogram(binwidth = 200)
```

```{r}
ggplot(penguins, aes(x = body_mass_g)) + 
        geom_histogram(binwidth = 20)
```

```{r}
ggplot(penguins, aes(x = body_mass_g)) + 
        geom_histogram(binwidth = 2000)
```

To check the skewness

```{r}
ggplot(penguins, aes(x = body_mass_g)) +
        geom_density()
```

## EXERCISES

1. Make a barplot of species of penguins, where you assign species to the y aesthetic. How is this plot different?

```{r}
ggplot(penguins, aes(y = species)) + 
        geom_bar()
```

2. How are the following two plots different? Which aesthetic, color or fill, is more useful for changing the color of bars?

```{r}
ggplot(penguins, aes(x = species)) +
        geom_bar(color = "red")
```

```{r}
ggplot(penguins, aes(x = species)) +
        geom_bar(fill = "red")
```

3. What does the bins argument in geom_histogram() do?

Agrupa mas o menos valores por intervalos. Por ejemplo podemos indicar 20 bins por lo que el histograma se nos agrupara por bloques de un intervalo de 20 en 20 del mismo tamaño. Si aumentamos el numero de bins tendremos menos barras porque los intervalos seran más grandes. En cambio si disminuimos los bins estaremos disminuyendo los intervalos y por lo general tendremos más barras.


4. Make a histogram of the carat variable in the diamonds dataset that is available when you load the tidyverse package. Experiments with different binwith values. What value reveals the most interesting patterns?

```{r}
ggplot(diamonds, aes(x = carat)) + 
        geom_histogram(binwidth = 10)
```

```{r}
ggplot(diamonds, aes(x = carat)) + 
        geom_histogram(binwidth = 1)
```

```{r}
ggplot(diamonds, aes(x = carat)) + 
        geom_histogram(binwidth = 0.1)
```


## VISUALIZING RELATIONSHIPS
## A numerical and a categorical variable

```{r}
ggplot(penguins, aes(x = species, y = body_mass_g)) +
        geom_boxplot()
```

```{r}
ggplot(penguins, aes(x = body_mass_g, color = species)) +
        geom_density(linewith = 0.75)
```

```{r}
ggplot(penguins, aes(x = body_mass_g, color = species, fill = species)) +
        geom_density(alpha = 0.5)
```

## Two categorical variables

```{r}
ggplot(penguins, aes(x= island, fill = species)) +
        geom_bar()
```

```{r}
ggplot(penguins, aes(x= island, fill = species)) +
        geom_bar(position = "fill")
```

## Two numerical variables

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
        geom_point()
```

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
        geom_point(aes(color = species, shape = island))
```

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
        geom_point(aes(color = species, shape = island)) +
        facet_wrap(~island)
```

